---
title: "tidyhydat intro"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

```{r, eval = FALSE}
help(package = "tidyhydat")
?hy_daily_levels
```

## Steps for tidyhydat
### Your turn
- Try grabbing daily flow data for the following stations: 07EF001, 02HA013
- Give each resulting object a unique name
```{r}
peace <- hy_daily_flows(station_number = "07EF001")

niagara <- hy_daily_flows(station_number = "02HA013")
```

- Inspect the results
- In the environment pane, click `niagara` and `peace` to view each object. `hy_daily_flows()` creates a ***data.frame object***

## Creating vectors
In R, you can use `c()` to *combine* other objects or values. 

```{r, create_vector}
sams_vector <- c("this","is","my","vector")

sams_vector2 <- c(1, 4, 5, 6)
```

### Your turn 
Station numbers can be supplied directly to the function like this:
```{r}
hy_daily_flows(station_number = c("07EF001","02HA013"))
```

Create a vector using `c()` called `my_vector`:
```{r}
## Your turn
```

Construct a data.frame object with `hy_daily_flows` that includes both stations
```{r}

```


Inspect that data.frame 
```{r}

```




