---
title: "tidyhydat intro"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```


## Your turn 1
Extract daily flow data for two stations and give each a unique name

```{r}
peace <- hy_daily_flows(station_number = "07EF001")

niagara <- hy_daily_flows(station_number = "02HA013")
```

In the environment pane, click `niagara` and `peace` to view each object. `hy_daily_flows()` creates a ***data.frame object***

## Your turn 2
In R, you can use `c()` to *combine* other objects or values. 

```{r, create_vector}
sams_vector <- c("this","is","my","vector")

sams_vector2 <- c(1, 4, 5, 6)
```

Station numbers can be supplied directly to the function like this:
```{r}
hy_daily_flows(station_number = c("07EF001","02HA013"))
```

Or they can be created initially as a separate vector which is supplied to the function:
```{r}
## Your turn
```

To find all the functions in a package we can input this line:
```{r, eval = FALSE}
help(package = "tidyhydat")
```

To seek out help for an individual function type `?function_name`. For example:
```{r, eval = FALSE}
?hy_daily_levels
```



