---
title: "Transform data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/hadley/data-science-in-tidyverse 
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

## hy_stn_data_range

```{r}
data_range <- hy_stn_data_range()
```

## filter()

```{r}
filter(data_range, DATA_TYPE == "Q")
```

## Your Turn

Use the logical operators to show:

Stations that begun operation before and including the year 1900
```{r}
filter(data_range)
```

Stations that are not discharge stations (i.e. "Q")
```{r}
filter(data_range)
```

Stations that have more than 100 years of data
```{r}
filter(data_range)
```

## Your Turn 

Use Boolean operators to return only the rows that contain:

* Find stations that are level ("H") and have more than 80 years record

```{r}
filter(data_range)
```

*  Find stations that start before 1890 or after 2016

```{r}
filter(data_range)
```

* These stations in one data.frame 08KA009, 05JA005, 06AC006 (hint look at %in%)

```{r}
filter(data_range)
```


## Quiz 

A function that returns a vector the same length as the input is called **vectorized**.

Which of the following functions are vectorized?

  * `ifelse()`
  * `diff()`
  * `sum()`

You might try these:
```{r}
data_range %>% 
  mutate(size = ifelse(pop < 10e06, "small", "large"))
```

```{r, error = TRUE}
data_range %>% 
  mutate(diff_pop = diff(pop))
```

```{r}
data_range %>% 
  mutate(total_pop = sum(as.numeric(pop)))
```

## Your Turn 5

Alter the code to add a `prev_lifeExp` column that contains the life expectancy from the previous record.

(Hint: use cheatsheet, you want to offset elements by one)

**Challenge**: Why isn't this quite the 'life expectency five years ago'?

```{r}
data_range %>%
  mutate()
```

## Your Turn 6

Use summarise() to compute three statistics about the data:

* The first (minimum) year in the dataset
* The last (maximum) year in the dataset
* The number of countries represented in the data (Hint: use cheatsheet)

```{r}
data_range 
```

## Your Turn 7

Extract the rows where `continent == "Africa"` and `year == 2007`. 

Then use summarise() and summary functions to find:

1. The number of unique countries
2. The median life expectancy

```{r}
data_range 
```

## Your Turn 8

Find the median life expectancy by continent in 2007.

```{r}
data_range %>%
  filter(year == 2007)
```

## Your Turn 9

Brainstorm with your neighbor the sequence of operations to find the country with biggest jump in life expectancy (between any two consecutive records) for each continent.

## Your Turn 10

Find the country with biggest jump in life expectancy (between any two consecutive records) for each continent.

```{r}


```

## Your Turn 11

Use `left_join()` to add the country codes in `country_codes` to the data_range data.

```{r}
country_codes
```

**Challenge**: Which codes in country_codes have no matches in data_range?

```{r}

```


***

# Take aways

* Extract cases with `filter()`  
* Make new variables, with `mutate()`  
* Make tables of summaries with `summarise()`  
* Do groupwise operations with `group_by()`
* Connect operations with `%>%`  
* Joins are two table verbs
