---
title: "Transform data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/hadley/data-science-in-tidyverse 
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

## hy_stn_data_range()
When `STATION_NUMBER` and `PROV_TERR_STATE_LOC` is missing all stations in Canada are returned. 

```{r}
data_range <- hy_stn_data_range()
```

## filter()

```{r}
filter(data_range, DATA_TYPE == "Q")
```

## Your Turn

Use the logical operators to show:

Stations that begun operation before and including the year 1900
```{r}
filter(data_range)
```

Stations that are not discharge stations (i.e. "Q")
```{r}
filter(data_range)
```

Stations that have more than 100 years of data
```{r}
filter(data_range)
```

## Your Turn 

Use Boolean operators to return only the rows that contain:

* Find stations that are level ("H") and have more than 80 years record

```{r}
filter(data_range)
```

*  Find stations that start before 1890 or after 2016

```{r}
filter(data_range)
```

* These stations in one data.frame 08KA009, 05JA005, 06AC006 (hint look at `%in%`)

```{r}
filter(data_range)
```

## pull()
`pull()` is a very simple function but as we will see it will help us feed the station numbers we want to the data that we need.

### Your turn
```{r}
##YEAR_FROM
pull(data_range)

##STATION_NUMBER
pull(data_range)
```

## Pipe `%>%` 

These do the same thing. Try it:
```{r}
filter(data_range, DATA_TYPE == "Q")
data_range %>% filter(DATA_TYPE == "Q")
```

## mutate()
```{r}
hy_monthly_flows("08MF005") %>% mutate(Value_cfs = Value = 0.0167)
```

```{r}
x <- 1:10
ifelse(x > 5, "large", "small")
```

### Your turn 
```{r}
hy_monthly_flows("08MF005") %>% 
  mutate(category = ifelse(Value > 500, "Large", "Small"))
```


First we need some data to work with:
```{r}
annual_gt_100 <- hy_stn_data_range() %>% 
   filter(RECORD_LENGTH > 100, DATA_TYPE == "Q") %>%                                          
   pull(STATION_NUMBER) %>%
   hy_annual_stats()
```

First we can see `mean()` with `na.rm = FALSE`, the default

```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value))
```

Note the difference with when we set `na.rm = TRUE`
```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value, na.rm = TRUE))
```

We can also compute more summaries:

```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value, na.rm = TRUE), 
            min_value = min(Value, na.rm = TRUE))
```

### Your turn

Use `summarise()` and the `annual_gt_100 data to compute three statistics about the data:

* The first (minimum) year in the dataset
* The median mean annual flow (Value) in the dataset
* The number of stations represented in the data (Hint: use cheatsheet)

```{r}
annual_gt_100 %>%
  summarise()
```

### Your Turn

Compute the summaries as before but only the mean annual flow (i.e. where Sum_stat is MEAN)

```{r}


```

## `group_by()`

```{r}
annual_gt_100 %>% 
   group_by(STATION_NUMBER)
```

Summarise by group
```{r}
annual_gt_100 %>% 
   group_by(Year) %>%
   summarise(num_stations = n_distinct(STATION_NUMBER))
```

### Your Turn 

Find the median life expectancy by continent in 2007.

```{r}
data_range %>%
  filter(year == 2007)
```

## Your Turn 9

Brainstorm with your neighbor the sequence of operations to find the country with biggest jump in life expectancy (between any two consecutive records) for each continent.

## Your Turn 10

Find the country with biggest jump in life expectancy (between any two consecutive records) for each continent.

```{r}


```

## Your Turn 11

Use `left_join()` to add the country codes in `country_codes` to the data_range data.

```{r}
country_codes
```

**Challenge**: Which codes in country_codes have no matches in data_range?

```{r}

```


***

# Take aways

* Extract cases with `filter()`  
* Make new variables, with `mutate()`  
* Make tables of summaries with `summarise()`  
* Do groupwise operations with `group_by()`
* Connect operations with `%>%`  
* Joins are two table verbs
