---
title: "Transform data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse 
-->

```{r packages}
library(tidyverse)
library(tidyhydat)
```

## hy_stn_data_range()
- When `STATION_NUMBER` and `PROV_TERR_STATE_LOC` is missing all stations in Canada are returned. 

```{r, eval = FALSE}
?hy_stn_data_range
```

Create an object from hy_stn_data_range()  called data_range which omit all arguments
```{r}
data_range <- hy_stn_data_range()
```

## filter()

```{r}
filter(data_range, DATA_TYPE == "Q")
```

### Your turn

Use the logical operators to show:

- Stations that begun operation before and including the year 1900
```{r}
filter(data_range)
```

- Stations that are not discharge stations (i.e. `"Q"`)
```{r}
filter(data_range)
```

- Stations that have more than 100 years of data
```{r}
filter(data_range)
```

### Your turn 

Use Boolean operators to return only the rows that contain:

- Find stations that are level (`"H"`) and have more than 80 years record

```{r}
filter(data_range)
```

- Find stations that start before 1890 or after 2016

```{r}
filter(data_range)
```

- These stations in one data.frame 08KA009, 05JA005, 06AC006 (hint look at `%in%`)

```{r}
filter(data_range)
```

### Two common mistakes
1. Use `=` instead of `==`
```{r error=TRUE}
filter(data_range, DATA_TYPE = "Q")
filter(data_range, DATA_TYPE == "Q")
```

2. Forgetting quotes
```{r error=TRUE}
filter(data_range, DATA_TYPE == Q)
filter(data_range, DATA_TYPE == "Q")
```


## pull()
`pull()` is a very simple function but as we will see it will help us feed the station numbers we want to the hydrometric data that we need.

### Your turn
```{r}
##YEAR_FROM
pull(data_range)

##STATION_NUMBER
pull(data_range)
```

## Pipe `%>%` 

These do the same thing. Try it:
```{r}
filter(data_range, DATA_TYPE == "Q")
data_range %>% filter(DATA_TYPE == "Q")
```

## mutate()
```{r}
hy_monthly_flows("08MF005") %>% mutate(Value_cfs = Value * 0.0167)
```

```{r}
x <- 1:10
ifelse(x > 5, "large", "small")
```

### Your turn 
```{r}
hy_monthly_flows("08MF005") %>% 
  mutate(category = ifelse(Value > 500, "Above 500", "Below 500"))
```

## summarise()
First we need some data to work with:
```{r}
annual_gt_100 <- hy_stn_data_range() %>% 
   filter(RECORD_LENGTH > 100, DATA_TYPE == "Q") %>%                                          
   pull(STATION_NUMBER) %>%
   hy_annual_stats()
```

First we can see `mean()` with `na.rm = FALSE`, the default

```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value))
```

Note the difference with when we set `na.rm = TRUE`
```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value, na.rm = TRUE))
```

We can also compute more summaries:

```{r}
annual_gt_100 %>%
  summarise(mean_value = mean(Value, na.rm = TRUE), 
            min_value = min(Value, na.rm = TRUE))
```

### Your turn

Use `summarise()` and the `annual_gt_100 data to compute three statistics about the data:

- The first (minimum) year in the dataset
- The median mean annual flow (Value) in the dataset
- The number of stations represented in the data (Hint: use cheatsheet)

```{r}
annual_gt_100 %>%
  summarise()
```


### Your Turn

Compute the summaries as before but only the mean annual flow (i.e. where Sum_stat is MEAN)

```{r}


```

## group_by()

```{r}
annual_gt_100 %>% 
   group_by(Year)
```

Summarise by group - the number of stations that have a record longer than 100 years grouped by year

```{r}
annual_gt_100 %>% 
   group_by(Year) %>%
   summarise(num_stations = n_distinct(STATION_NUMBER))
```

### Your Turn 

```{r}
annual_gt_100
```

## select
```{r}
annual_gt_100 %>% 
  select(STATION_NUMBER, Value)
```

## arrange
```{r}
annual_gt_100 %>% 
  arrange(Value)
```


## Challenge!
Extract daily flow information from all active stations on Prince Edward Island.
```{r}

```
