---
title: "Working with Dates and Joins"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/hadley/data-science-in-tidyverse 
-->


```{r setup}
library(tidyverse)
library(lubridate)
library(tidyhydat)
```

# Review
Remember some of the date types we have seen previously:

```{r}
fraser_flow <- hy_daily_levels(station_number = "08MF005")
```

# Times and Dates
## Your Turn 

For each of the following formats (of the same date), pick the right `ymd()` function to parse them:

```{r}
"2018 Feb 01"
"2-1-18"
"01/02/2018"
```

### Your Turn 

Fill in the blanks to for `fraser_flows` data:

- Extract the month from date.
- Extract the year from date.
- Calculate the mean flow with a measurement for each year/month.
- Filter the results for only the month of June


```{r, error = TRUE}
fraser_flow %>%
  mutate(year = _____,
         month = _____) %>%
  group_by(_____, ____) %>%
  summarise(mean_monthly_flow = _____) %>%
  filter(____)
```

### realtime data
```{r}
mackenzie_realtime <- realtime_dd(station_number = "10LC014")

mackenzie_realtime %>% 
mutate(seconds = second(Date), 
       minute = minute(Date),
       hour = hour(Date)) %>%   
filter(hour == 7)
```

or this way

```{r}
mackenzie_realtime %>% 
  filter(hour(Date) == 7) 
```

## filter() and dates
```{r}
"01-01-1950"

dmy("01-01-1950")
```


```{r}
fraser_flow %>%
  filter(Date >= "01-01-1950") %>%
  summarise(min_date = min(Date))
```

```{r}
fraser_flow %>%
  filter(Date >= dmy("01-01-1950"))%>%
  summarise(min_date = min(Date))
```


## Challenge

Get the province/territory station codes like this:
```{r}
unique(allstations$PROV_TERR_STATE_LOC)
```

Try using the `hy_stations` function for station metadata:
```{r}
hy_stations(prov_terr_state_loc = "BC")
```


# Joining data
```{r}
bc_stations <- hy_stations(prov_terr_state_loc = "BC") %>%
  select(STATION_NUMBER, STATION_NAME)

bc_mad <- hy_annual_stats(prov_terr_state_loc = "BC") %>%
  filter(Sum_stat == "MEAN")

bc_stations %>%
  left_join(bc_mad, by = c("STATION_NUMBER"))
```



## Your turn

Find the Unit flow (Flow per watershed area) for all BC stations for all years

```{r}

```

