---
title: "Working with Dates and Joins"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/hadley/data-science-in-tidyverse 
-->


```{r setup}
library(tidyverse)
library(lubridate)
```

# Warm-up / Review
Remember some of the date types we have seen previously:

```{r}
fraser_flow <- hy_daily_levels(station_number = "08MF005")
```

# Times and Dates
## Your Turn 

For each of the following formats (of the same date), pick the right `ymd()` function to parse them:

```{r}
"2018 Feb 01"
"2-1-18"
"01/02/2018"
```

### Your Turn 

Fill in the blanks to for `fraser_flows` data:

- Extract the month from date.
- Extract the year from date.
- Calculate the mean flow with a measurement for each year/month.
- Filter the results for only the month of June


```{r, error = TRUE}
fraser_flow %>%
  mutate(year = _____,
         month = _____) %>%
  group_by(_____, ____) %>%
  summarise(mean_monthly_flow = _____) %>%
  filter(____)
```

### realtime data
```{r}
mackenzie_realtime <- realtime_dd(station_number = "10LC014")

mackenzie_realtime %>% 
mutate(seconds = second(Date), 
       minute = minute(Date),
       hour = hour(Date)) %>%   
filter(hour == 7)
```

or this way

```{r}
mackenzie_realtime %>% 
  filter(hour(Date) == 7) 
```

## filter() and dates
```{r}
"01-01-1950"

dmy("01-01-1950")
```


```{r}
fraser_flow %>%
  filter(Date >= "01-01-1950")
```

```{r}
fraser_flow %>%
  filter(Date >= dmy("01-01-1950"))
```


# Challenge
```{r}

```


# Take Aways

Dplyr gives you three _general_ functions for manipulating data: `mutate()`, `summarise()`, and `group_by()`. Augment these with functions from the packages below, which focus on specific types of data.

Package   | Data Type
--------- | --------
forcats   | factors
stringr   | strings
hms       | times
lubridate | dates and times

# Joins



## Your Turn 11

Use `left_join()` to add the country codes in `country_codes` to the data_range data.

```{r}
country_codes
```

**Challenge**: Which codes in country_codes have no matches in data_range?

```{r}

```


***

# Take aways

* Extract cases with `filter()`  
* Make new variables, with `mutate()`  
* Make tables of summaries with `summarise()`  
* Do groupwise operations with `group_by()`
* Connect operations with `%>%`  
* Joins are two table verbs

