---
title: "Visualize data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse 
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

# ggplot2 and tidyhydat

Create a dataframe of the number of stations and average length of record by real time and activity status. 
```{r}
data_range <- hy_stn_data_range() %>%
  group_by(STATION_NUMBER) %>%
  summarise(total_record_length = sum(RECORD_LENGTH, na.rm = TRUE))

cdn_prov <- c("NB", "PE", "NS", "ON", "QC", "NL", "MB", "AB", "SK", "NU", "NT", "BC", "YT")

station_meta <- hy_stations() %>%  
  filter(PROV_TERR_STATE_LOC %in% cdn_prov) %>% 
  group_by(PROV_TERR_STATE_LOC, HYD_STATUS, REAL_TIME) %>%
  left_join(data_range, by = c("STATION_NUMBER")) %>% 
  summarise(number_stns = n(), mean_record_length = mean(total_record_length, na.rm = TRUE))
```

## Your turn

What relationships might interest you with this data? 

Type in the code from the slide, and run it to make a graph. Pay strict attention to spelling, capitalization, and parentheses.

I strongly recommend typing it, even though copying and pasting is easier! Re-typing it will help your fingers learn the patterns you'll need to apply to your own data analyses.

```{r}
ggplot(data = station_meta, mapping = aes(x = number_stns, y = mean_record_length)) +
  geom_point()
```

Remember to raise a post-it if you get stuck, and then once you're done!

# Mappings

## Your Turn 

Add `colour`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = station_meta, mapping = aes(x = number_stns, y = mean_record_length)) +
  geom_point()
```

# Geoms

## Your Turn
Use the mean annual flow from all PEI stations

```{r data}
pei_mean <- hy_annual_stats(prov_terr_state_loc = "PE") %>%
  filter(Sum_stat == "MEAN", Parameter == "Flow")
```

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Guess!

```{r}
ggplot(data = pei_mean) +
  geom_point(mapping = aes(x = STATION_NUMBER, y = Value)) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```

```{r boxplot}
ggplot(data = pei_mean)
```


## Your Turn

Make a hydrograph of the 08GB014 HORSESHOE RIVER ABOVE LOIS LAKE BC station using the Year and yearday columns created. For this we again need to make use of the lubridate package.

```{r}
library(lubridate)
horseshoe <- hy_daily_flows(station_number = "08GB014") %>% 
  mutate(Year = year(Date), yearday = yday(Date))
```

### Grouping
```{r group}
ggplot(data = horseshoe)
```

### Facet
```{r facet}
ggplot(data = horseshoe)
```

## Your Turn

Make a histogram plot of `horseshoe`.

```{r}
ggplot(data = horseshoe, mapping = aes(x = Value)) +
  geom_histogram()
```

And then if you would like to modify the bin size, simply supply the bidwidth argument. Experiment with different values:
```{r histo}
ggplot(data = horseshoe, mapping = aes(x = Value)) +
  geom_histogram(binwidth = 1) 
```

And to add your distribution visualization, we can add *rug* marks. Note that the colour aesthetic is only mapped to the `geom_rug` function:

```{r rug}
ggplot(data = horseshoe, mapping = aes(x = Value)) +
  geom_histogram(binwidth = 1) +
  geom_rug(mapping = aes(colour = Value))
```

# Saving Graphs

## Your Turn

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane. Next find it on your desktop.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

# What else?

## Basic plot
```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS))
```

## Titles and captions + labs()
```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS)) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory")
```

## Modifying the legend title
```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS)) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_discrete(name = "Status")
```

## Position Adjustments

```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS), position = position_dodge()) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_discrete(name = "Status")
```

## Theme - Visual appearance of non-data elements

```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS), position = position_dodge()) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_discrete(name = "Status") +
  theme_minimal()
```

## Scales - Customize color scales, other mappings

```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS), position = position_dodge()) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_brewer(name = "Status", palette = "Spectral") +
  theme_minimal()
```

## Facets - Subplots that display subsets of the data.

```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS), position = position_dodge()) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_brewer(name = "Status", palette = "Spectral") +
  theme_minimal() +
  facet_wrap(~REAL_TIME, ncol = 1)
```

## Coordinate systems

```{r}
ggplot(data = station_meta, mapping = aes(x = PROV_TERR_STATE_LOC, y = number_stns)) +
  geom_bar(stat = "identity", aes(fill = HYD_STATUS), position = position_dodge()) +
  labs(title = "Hydrometric Stations in Canada",
       subtitle = "All stations are part of the Water Survey of Canada network",
       caption = "Figured generated with tidyhydat and ggplot2",
       y = "Number of stations",
       x = "Province /Territory") +
  scale_fill_brewer(name = "Status", palette = "Spectral") +
  theme_minimal() +
  facet_wrap(~REAL_TIME) +
  coord_polar()
```