---
title: "Visualize data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse 
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

The number of stations and average length of record by real time and activity status. 
```{r}
data_range <- hy_stn_data_range() %>%
  group_by(STATION_NUMBER) %>%
  summarise(total_record_length = sum(RECORD_LENGTH, na.rm = TRUE))

cdn_prov <- c("NB", "PE", "NS", "ON", "QC", "NL", "MB", "AB", "SK", "NU", "NT", "BC", "YT")

station_meta <- hy_stations() %>%  
  filter(PROV_TERR_STATE_LOC %in% cdn_prov) %>% 
  group_by(PROV_TERR_STATE_LOC, HYD_STATUS, REAL_TIME) %>%
  left_join(data_range, by = c("STATION_NUMBER")) %>% 
  summarise(number_stns = n(), mean_record_length = mean(total_record_length, na.rm = TRUE))
```

## Your turn

What relationships might interest you with this data? 

Type in the code from the slide, and run it to make a graph. Pay strict attention to spelling, capitalization, and parentheses.

I strongly recommend typing it, even though copying and pasting is easier! Re-typing it will help your fingers learn the patterns you'll need to apply to your own data analyses.

```{r}
ggplot(data = station_meta, mapping = aes(x = number_stns, y = mean_record_length)) +
  geom_point()
```

Remember to raise a post-it if you get stuck, and then once you're done!

## Your Turn 2

Add `colour`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = station_meta, mapping = aes(x = number_stns, y = mean_record_length)) +
  geom_point()
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Guess!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 5

Make a density plot of `hwy` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 6

Make a bar chart `class` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane. Next find it on your desktop.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```
