---
title: "Visualize data"
---

<!-- 
This file Creative Commons Attribution 4.0 International License. 
See details at https://github.com/bcgov/intro-to-tidyhydat-and-tidyverse 
-->

```{r setup}
library(tidyverse)
library(tidyhydat)
```

Our data
```{r}
pei_stations <- hy_stations(prov_terr_state_loc = "PE") %>%
  filter(HYD_STATUS == "ACTIVE") %>% 
  select(STATION_NUMBER, STATION_NAME, DRAINAGE_AREA_GROSS)

pei_mad <- hy_annual_stats(prov_terr_state_loc = "PE") %>%
  filter(Sum_stat == "MEAN") %>%
  left_join(pei_stations, by = c("STATION_NUMBER")) %>%
  select(STATION_NUMBER, STATION_NAME, Parameter, Year, Value, DRAINAGE_AREA_GROSS)

pei_mad
```

## Your turn

What relationship do you expect to see between drainage area (DRAINAGE_AREA_GROSS) and mean annual flow (Value)?

Type in the code from the slide, and run it to make a graph. Pay strict attention to spelling, capitalization, and parentheses.

I strongly recommend typing it, even though copying and pasting is easier! Re-typing it will help your fingers learn the patterns you'll need to apply to your own data analyses.

```{r}
ggplot(data = pei_mad, aes(x = DRAINAGE_AREA_GROSS, y = Value)) +
  geom_point() 
```

Remember to raise a post-it if you get stuck, and then once you're done!

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Guess!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 5

Make a density plot of `hwy` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 6

Make a bar chart `class` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane. Next find it on your desktop.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```
